### 2021-02-20 mathematics.tex
more on spinors

\paragraph{Clifford algebra and spin group}\TODO{needs experts...}
We just provide a little information on spin group $\mathop{\mathrm Spin}(1,3)$ associated to (our) Minkowski metric $\eta=\diag(+,-,-,-)$:~\cite{RauschdeTraubenberg:2005aa,Wikipedia:Clifford_Algebra}. To define the Clifford algebra $\mathfrak C(V,Q)$, let
\begin{itemize}
 \item[$V$]: a vector space over a field $K$; in our case, $V=\mathbb R^{1,3}$ and $K=\mathbb R$.
 \item[$Q$]: a quadratic form on $V$,
 \item[$A$]: %
  a unital associative algebra over $K$; it means that $(A, +, \ast)$ is a ring and is also a vector space on $K$, it should satisfy
  $(k a_1)\ast a_2=a_1\ast(ka_2)=k(a_1\ast a_2)$ for $a_i\in A$ and $k\in K$, and it should contain $1_A$ such that $1_A\ast a=a\ast 1_A=a$;
  here a map $f:K\to A$ with $f(k)=k1_A$ is an injective ring homomorphism, with which $K$ can be included in $A$.
\end{itemize}
Then we can define a linear map $i:V\to \mathfrak C(V,Q)$ satisfying $i(v)\ast i(v)=Q(v)1_A$ for all $v\in V$ as follows, which defines the Clifford algebra.
Given a linear map $j:V\to A$ satisfying $j(v)\ast j(v)=Q(v)1_A$ for all $v\in V$, there is a unique algebra homomorphism $f:\mathfrak C(V,Q)\to A$ such that $f\circ i=j$.

>>>>

\paragraph{Clifford algebra and spin group}\TODO{needs experts...}
We just provide a little information on spin group $\mathop{\mathrm Spin}(1,3)$ associated to (our) Minkowski metric $\eta=\diag(+,-,-,-)$:~\cite{RauschdeTraubenberg:2005aa,Yamaguchi:spinor,Wikipedia:Clifford_Algebra}. To define the Clifford algebra $\mathfrak C(V,Q)$, let
\begin{itemize}
 \item[$V$]: a vector space over a field $K$; in our case, $V=\mathbb R^{1,3}$ and $K=\mathbb R$.
 \item[$Q$]: a quadratic form on $V$,
 \item[$A$]: %
  a unital associative algebra over $K$; it means that $(A, +, \ast)$ is a ring and is also a vector space on $K$, it should satisfy
  $(k a_1)\ast a_2=a_1\ast(ka_2)=k(a_1\ast a_2)$ for $a_i\in A$ and $k\in K$, and it should contain $1_A$ such that $1_A\ast a=a\ast 1_A=a$;
  here a map $f:K\to A$ with $f(k)=k1_A$ is an injective ring homomorphism, with which $K$ can be included in $A$.
\end{itemize}
Then we can define a linear map $i:V\to \mathfrak C(V,Q)$ satisfying $i(v)\ast i(v)=Q(v)1_A$ for all $v\in V$ as follows, which defines the Clifford algebra.
Given a linear map $j:V\to A$ satisfying $j(v)\ast j(v)=Q(v)1_A$ for all $v\in V$, there is a unique algebra homomorphism $f:\mathfrak C(V,Q)\to A$ such that $f\circ i=j$.

### 2021-02-20 spinors.tex
more on spinors

\changefontsizes{10pt}
\section{Spinors}

\changefontsizes{9pt}
\begin{equation}
 (\overline{\psi_1}\psi_2)^* = (\psi_2)^\dagger(\overline\psi_1)^\dagger = \overline{\psi_2}\psi_1.
\end{equation}

@@

\paragraph{More isomorphism}
Let us see a bit more of  mathematical structure, following the discussion in \cref{sec:group-theory}.
The isomorphic groups
$\mathop{\mathrm{Spin}}(1,3)^+\cong\gSL(2,\mathbb C)\cong\gSp(2,\mathbb C)$
is a double cover of $L_0$;
in particular, $L_0\cong \mathop{\mathrm{PSL}}(2;\mathbb C)=\gSL(2;\mathbb C)/\mathbb Z_2$.

@@

\paragraph{Spinors from Clifford algebra}
First we construct spinors from Clifford algebra $\mathfrak C_{1,3}$ given in \cref{sec:group-theory}. A matrix representation $\gamma^\mu$ of $\mathfrak C_{1,3}$ and spinors $\psi$ are introduced:
\begin{equation}
 \overline\psi = \psi^\dagger\beta, \quad\beta\beta=1,\qquad
\{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu}, \quad
\gamma_5 = h_5\times\ii\gamma^0\gamma^1\gamma^2\gamma^3; \qquad
\psi\mapsto T\psi,\quad
\overline\psi\mapsto \overline\psi\beta T^\dagger \beta;\quad T\in \Gamma_{1,3}.
\end{equation}
Thus, in order that $\overline\psi\psi$ and $\overline\psi\gamma^\mu\psi$ be respectively scalar and vector, $T$ should satisfy
\begin{equation}
 T^{-1}\gamma^\mu T = \LorTr\mu\nu\gamma^\nu,\quad
 \beta T^\dagger \beta T=1,
\end{equation}
or in infinitesimal form $T=1+(-\ii/2)d^{\mu\nu}S_{\mu\nu}$,
\begin{equation}
 (S_{\mu\nu})^\dagger = \beta S_{\mu\nu} \beta,\quad
[S_{\mu\nu},\gamma^\alpha] = -(M_{\mu\nu})^{\alpha}{}_\beta\gamma^\beta;
\qquad
\therefore S_{\mu\nu}=\frac{\ii}{4}[\gamma_\mu,\gamma_\nu];
\quad
[\gamma_\mu^\dagger, \gamma_\nu^\dagger]=\beta[\gamma_\mu,\gamma_\nu]\beta;
\end{equation}
finally, a representation of the Lorentz group is constructed as
\begin{equation}
\Lambda\stackrel{\text{rep}}=
 \exp\left(\frac{-\ii}{2}d^{\mu\nu}S_{\mu\nu}\right);\qquad
 [S_{\mu\nu},S_{\rho\sigma}] = -\ii\left(
\eta_{\mu\rho} S_{\nu\sigma}
-\eta_{\nu\rho} S_{\mu\sigma}
+\eta_{\mu\sigma} S_{\nu\rho}
-\eta_{\nu\sigma} S_{\mu\rho}\right),
\end{equation}
which is exactly the same as \cref{eq:LorentzAlgebra}.

>>>>

\summarystyle

\section{Spinors}
\begin{alignat}{2}
&\text{Gamma matrices}:&\quad&
\{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu},~
\gamma_5=\ii\gamma^0\gamma^1\gamma^2\gamma^3;\quad
\{\gamma^\mu,\gamma_5\}=0,~
\gamma^5\gamma^5=1.
\\
&\text{conjugates}:&&
\overline{\psi}=\psi^\dagger \beta,\qquad
\psi^\cc=C(\overline\psi)^\TT
\end{alignat}

\paragraph{chiral notation}
\begin{align}\label{eq:chiralnotation}
& \overline{\psi}=\psi^\dagger\gamma^0;~
 \gamma^\mu=\pmat{0&\sigma^\mu\\\bar\sigma^\mu&0},~
 \gamma_5=\pmat{-1&0\\0&1};~
 \PL = \frac{1-\gamma_5}{2},~
 \PR = \frac{1+\gamma_5}{2}.
\\&
(\gamma^\mu)^\dagger = \gamma^0\gamma^\mu\gamma^0,~
(\gamma^\mu)^*       = \gamma^2\gamma^\mu\gamma^2,~
(\gamma^\mu)^\TT     = \gamma^0\gamma^2\gamma^\mu\gamma^2\gamma^0,
\end{align}

\paragraph{(cf.~Dirac notation)}
\begin{align}\label{eq:diracnotation}
& %\overline{\psi}=\psi^\dagger\hat\gamma^0;~
 \hat\gamma^0=\pmat{1&0\\0&-1},~
 \hat\gamma^i=\pmat{0&\sigma^i\\\bar\sigma^i&0},~
 \hat\gamma_5=\pmat{0&1\\1&0};~
 \hat\PL = \frac{1-\hat\gamma_5}{2},~
 \hat\PR = \frac{1+\hat\gamma_5}{2}.
\\&
(\hat\gamma^\mu)^\dagger = \hat\gamma^0\hat\gamma^\mu\hat\gamma^0,~
(\hat\gamma^\mu)^*       = \hat\gamma^2\hat\gamma^\mu\hat\gamma^2,~
(\hat\gamma^\mu)^\TT     = \hat\gamma^0\hat\gamma^2\hat\gamma^\mu\hat\gamma^2\hat\gamma^0,
\end{align}

\begin{equation}
 (\overline{\psi_1}\psi_2)^* = (\psi_2)^\dagger(\overline\psi_1)^\dagger = \overline{\psi_2}\psi_1.
\end{equation}

\newpage
\detailstyle

@@

\paragraph{More isomorphism}
Let us see a bit more of  mathematical structure, following the discussion in \cref{sec:group-theory} (cf.~Refs.~\cite{RauschdeTraubenberg:2005aa,Yamaguchi:spinor}).
The isomorphic groups
$\mathop{\mathrm{Spin}}(1,3)^+\cong\gSL(2,\mathbb C)\cong\gSp(2,\mathbb C)$
is a double cover of $L_0$;
in particular, $L_0\cong \mathop{\mathrm{PSL}}(2;\mathbb C)=\gSL(2;\mathbb C)/\mathbb Z_2$.

@@

\paragraph{Representation of Clifford algebra}
As summarized in \cref{sec:group-theory}, the spin group $\mathop{\mathrm{Spin}}(1,3)^+$ generated by Clifford algebra $\mathfrak C_{1,3}$ is a double cover of $L_0$, and thus we can consider a representation of $L_0$ based on $\mathfrak C_{1,3}$.

To construct an irreducible representation of $\mathfrak C_{1,3}$, we utilize the fact that we can form two sets of creation-annihilation operators
\begin{align}
 a^\pm = \sqrt{h_\eta}\frac{e^0\pm e^3}{2},\quad
 b^\pm = \sqrt{h_\eta}\frac{\pm e^2-\ii e^1}{2};\qquad
 \{a^+,a^-\}=1,~
 \{b^+,b^-\}=1,~
 \{\text{(others)}\} = 0.
\end{align}
These ladder operator allows us to construct four states starting from $\ket{00}$, which is a non-zero state with $a^-\ket{00}=b^-\ket{00}=0$, and to construct an irreducible representation of $\mathfrak C_{1,3}$ (and, in fact, it is unique for even dimension):
\begin{equation}
 \ket{10} = a^+\ket{00},\quad
 \ket{01} = b^+\ket{00},\quad
 \ket{11} = a^+b^+\ket{00}\quad\leadsto\quad
 a^+=\spmat{0&0&1&0\\0&0&0&1\\0&0&0&0\\0&0&0&0},\quad
 b^+=\spmat{0&-1&0&0\\0&0&0&0\\0&0&0&1\\0&0&0&0},
\end{equation}
and $a^-=(a^+)^\dagger$, $b^-=(b^+)^\dagger$.
We then obtain a representation $\gamma$, which is called ``standard representation.''\footnote{The chiral notation \eqref{eq:chiralnotation} and Dirac notation \eqref{eq:diracnotation} are equivalent to this standard representation, i.e., related by unitary matrices.}
They are not Hermitian, but as we will see, this non-Hermiticity is solved by amending the inner product by a matrix $\beta$: $(\psi,\gamma^\mu\psi):=\psi^\dagger\beta \gamma^\mu\psi$.

Although $\psi$ forms an irreducible representation $\gamma^\mu$ of $\mathfrak C_{1,3}$, the resulting representation $S_{\mu\nu}$ (see the next paragraph) is a reducible representation of $\mathop{\mathrm {Spin}}(1,3)^+$.
This is confirmed by
\begin{equation}
 \gamma_5\gamma_5=1,\quad
\{\gamma_5,\gamma^\mu\}=0,\quad
 [\gamma_5,S_{\mu\nu}]=0;\qquad \gamma_5:=\ii\gamma^0\gamma^1\gamma^2\gamma^3,
\end{equation}
and $P{}^{\text L}\w{R} =(1\mp\gamma_5)/2$ works as the projection operators.
In addition, four state are eigenstates of $J_3=h_\eta S_{12}$ because 
\begin{equation}
 [J_3,b^+b^-] = 0,~
 J_3=b^+b^--\frac12,
\end{equation}
which also guarantees that spinors have spin $1/2$.
In summary,
\begin{align}
 &\ket{00} =\ket{-\w L},\quad
 \ket{10} =:\ket{-\w R},\quad
 \ket{01} =:\ket{+\w R},\quad
 \ket{11} =:\ket{+\w L};\\
&J_3\ket{\pm\w H}=\pm\tfrac12\ket{\pm\w H},\quad
\PL\ket{\pm\w L}=\ket{\pm\w L},~
\PR\ket{\pm\w R}=\ket{\pm\w R};\quad
\PL\ket{\pm\w R}=
\PR\ket{\pm\w L}=0.
\end{align}
For example, in chiral notation with $(+,-,-,-)$, the Lorentz generators $S_{\mu\nu}$ are block diagonal and $\ket{\pm\w L}$ ($\ket{\pm\w R}$) has non-zero component only in the upper (lower) two component:
\begin{equation}
 \ket{-\w L} = \spmat{0\\1\\0\\0},~
 \ket{-\w R} = \spmat{0\\0\\0\\1},~
 \ket{+\w R} = \spmat{0\\0\\\ii\\0},~
 \ket{+\w L} = \spmat{\ii\\0\\0\\0}.
\end{equation}

\paragraph{Four-spinors and Lorentz transformation}
The above ``theoretical'' discussion can be seen more explicitly, starting from spinors and a matrix representation $\gamma^\mu$ given by
\begin{equation}
 \overline\psi = \psi^\dagger\beta, \quad\beta\beta=1,\qquad
\{\gamma^\mu,\gamma^\nu\}=2\eta^{\mu\nu};\qquad
\psi\mapsto T\psi,\quad
\overline\psi\mapsto \overline\psi\beta T^\dagger \beta;\quad T\in \mathop{\mathrm{Spin}}(1,3)^+.
\end{equation}
For $\overline\psi\psi$ and $\overline\psi\gamma^\mu\psi$ to be respectively scalar and vector, $T$ should satisfy
\begin{equation}
 T^{-1}\gamma^\mu T = \LorTr\mu\nu\gamma^\nu,\quad
 \beta T^\dagger \beta T=1,
\end{equation}
or in infinitesimal form $T=1+(-\ii/2)d^{\mu\nu}S_{\mu\nu}$,
\begin{equation}
 (S_{\mu\nu})^\dagger = \beta S_{\mu\nu} \beta,\quad
[S_{\mu\nu},\gamma^\alpha] = -(M_{\mu\nu})^{\alpha}{}_\beta\gamma^\beta;
\qquad
\therefore S_{\mu\nu}=\frac{\ii}{4}[\gamma_\mu,\gamma_\nu];
\quad
[\gamma_\mu^\dagger, \gamma_\nu^\dagger]=\beta[\gamma_\mu,\gamma_\nu]\beta;
\label{eq:betacondition}
\end{equation}
the first condition leads to a representation of the Lorentz group
\begin{equation}
\Lambda\stackrel{\text{rep}}=
 \exp\left(\frac{-\ii}{2}d^{\mu\nu}S_{\mu\nu}\right);\qquad
 [S_{\mu\nu},S_{\rho\sigma}] = -\ii\left(
\eta_{\mu\rho} S_{\nu\sigma}
-\eta_{\nu\rho} S_{\mu\sigma}
+\eta_{\mu\sigma} S_{\nu\rho}
-\eta_{\nu\sigma} S_{\mu\rho}\right)
\end{equation}
as seen in \cref{eq:LorentzAlgebra}, while the second condition determines what $\beta$ should be, as given in, e.g., \cref{eq:chiralnotation} and \cref{eq:diracnotation}.

\paragraph{Charge conjugation and Majorana spinor}
The charge conjugation of $\psi$ is something like $\psi^*$ but should obey the same representation as $\psi$ does, i.e., $C'\psi^*$ with $C'$ being a unitary matrix such that $B\psi^*\to TB\psi^*$. Or it can be seen that the, the previous procedure with $\psi^*$ may generate another irreducible representation and it should be related to $\gamma^\mu$ by an unitary matrix. Anyway, we define the charge conjugation by
\begin{equation}
 \psi^\cc = C(\overline\psi)^\TT = C\beta^\TT\psi^*;\quad CC^\dagger = 1.
\quad
\therefore
C^*\beta[\gamma_\mu^\dagger,\gamma_\nu^\dagger]\beta C^\TT =
[\gamma_\mu^\TT,\gamma_\nu^\TT].
\end{equation}
Combining with \cref{eq:betacondition},
\begin{equation}
 \beta\beta=1,\quad C C^\dagger = 1,\quad
 \beta\gamma^\mu\beta=h_\beta(\gamma^\mu)^\dagger,\quad
 C\gamma_\mu^* C^\dagger = h_C\gamma_\mu^\dagger.
\end{equation}
In even dimensions, the expressions have many choices as seen in the signs $h_C$ and $h_\beta$; moreover, the sign $h\w{cc}$ defined in $(\psi^\cc)^\cc=h\w{cc}\psi$ depends on the definition.
It is thus useful to use a specific notation for further discussion.

For example, the construction of a Majorana spinor $\psi\w M$, which satisfies $(\psi\w M)^\cc=\psi\w M$, is simply done as $\psi\w{M}\propto\psi+\psi^\cc$ if $\eta\w{cc}=1$, but needs some phases if $\eta\w{cc}=-1$.

\paragraph{Weyl spinor}
